---
title: "Data Wrangling"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(gridExtra)
theme_set(theme_light())
```

```{r message=FALSE}
# Read in our CSV's
train <- read_csv("train.csv")
test <- read_csv("test.csv")

# Make tibbles equal for rbind and bind them together
data <- bind_rows(
  train %>% 
    mutate(tibble = "train"),
  test %>% 
    mutate(Survived = as.integer(""), 
           tibble = "test"))
```

### What are we missing?

Let's right away get a sense of our missing data.

```{r}
data %>% 
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))
```


### Names

Full names in a single column are never a good idea in data. Why don't we organize that column?

```{r}
# Name column split on '.'
data$Name %>% 
  str_split(pattern = "\\.", simplify = TRUE)

# Name column split on ','
data$Name %>% 
  str_split(pattern = ",", simplify = TRUE)

# Last name separation seems to be more consistent
# Separate LastName first
data <- data %>% 
  separate(col = Name, into = c("LastName", "TitleFirstMiddle"), sep = ",")

# Modify the single instance where other name includes '.'
data$TitleFirstMiddle <- data$TitleFirstMiddle %>% 
  str_replace(pattern = "Elizabeth L. Barrett", replacement = "Elizabeth L Barrett")

# Separate Title from First, Middle and other
data <- data %>% 
  separate(col = TitleFirstMiddle, into = c("Title", "FirstMiddleOther"), sep = "\\.")

# Modify 'The Countess' name and title
data$LastName <- data$LastName %>% 
  str_replace(pattern = "Rothes", replacement = "Dyer-Edwards")
data$Title <- data$Title %>% 
  str_replace(pattern = " the Countess", replacement = "The Countess of Rothes")
data$FirstMiddleOther <- data$FirstMiddleOther %>% 
  str_replace(pattern = " of \\(Lucy Noel Martha Dyer-Edwards\\)", replacement = "Lucy Noel Martha")

# Remove beginning empty space
data$FirstMiddleOther <- data$FirstMiddleOther %>% str_replace(pattern = "^ ", replacement = "")
data$Title <- data$Title %>% str_replace(pattern = "^ ", replacement = "")

# Let's include a binary variable that indicates whether someone has another name
data$OtherName <- str_detect(data$FirstMiddleOther, pattern = "\\(")
```

### Tickets

Let's examine what `Ticket` means and how we can derive information from it.

```{r}
data %>% 
  group_by(Ticket) %>% 
  tally(sort = TRUE)

data %>% 
  filter(Ticket=="1601")

# Lets separate the ticket letters and numbers,
# but first we need to clear the spaces and other characters
data$Ticket <- str_replace_all(string = data$Ticket, pattern = "[^[:alnum:]]", replacement = "")

data$TicketNumber <- as.numeric(str_extract(data$Ticket, "[0-9]+"))
data$TicketLetter <- str_extract(data$Ticket, "[aA-zZ]+")

# We've found that four 'LINE' tickets were issued, so let's give them ticket #0
data %>% 
  filter(is.na(TicketNumber))

data$TicketNumber <- replace_na(data = data$TicketNumber, replace = 0)

# We could populate all missing values of TicketLetter with unknown, 
# but it really wouldn't serve any meaningful purpose
#data$TicketLetter %>% replace_na(replace = "UNKNOWN")

data %>% group_by(TicketLetter) %>% tally(sort = TRUE)

# Are there different column values for one ticket?
data %>% 
  group_by(Ticket) %>% 
  summarise(dis = n_distinct(LastName)) %>% 
  arrange(desc(dis))
```

## Filling in Missing Data

There is only one **fare** missing.

```{r missing fare}
# Which row(s) contain missing data
data %>% 
  filter(is.na(Fare))

# Let's find similar values to the missing one
data %>% 
  filter(Pclass==3,
         Sex=="male",
         Embarked=="S",
         Age>60)

# Another passenger is very similar to we will use that fare
data$Fare <- data$Fare %>% 
  replace(is.na(.), 6.2375)
```

**Embarked** has 2 missing values. Let's take care of that now.

```{r missing embarked}
# Which row(s) contain missing data
data %>% 
  filter(is.na(Embarked))

# Let's find similar values to the missing two
data %>% 
  filter(Pclass==1,
         Sex=="female",
         Title=="Miss",
         grepl('B', Cabin))

# Visualize the likelihood of 'Port of Embarkation' against other variables
data %>% ggplot(aes(Embarked, Cabin)) + geom_jitter()

# We can rule out Q quite easily
# How about others
data %>% ggplot(aes(Embarked, Pclass)) + geom_jitter()

# Looking at similar people, 'S' seems most suited for both missing values
data$Embarked <- data$Embarked %>% 
  replace(is.na(.), "S")
```

**Age** has 263 missing vlaues. Let's get those filled with the most suited values.

```{r missing age}
# Which row(s) contain missing data
data %>% 
  filter(is.na(Age))

# They are certainly too many to fill in, 
# we have to develop a clever solution.
# Filling in the missing values with the broadest selection of groups
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data <- data %>%
    group_by(Pclass, Title, Sex, SibSp, Parch, Embarked) %>%
    mutate(
        Age = impute.mean(Age)
    ) %>% ungroup()

# We only have 28 NA's for Age left, let's reduce the number of groupings

data <- data %>%
    group_by(Pclass, Title, Sex, Parch, Embarked) %>%
    mutate(
        Age = impute.mean(Age)
    ) %>% ungroup() 

# Now we have only 15 left, let's do something similar

data <- data %>%
    group_by(Pclass, Title, Sex, Embarked) %>%
    mutate(
        Age = impute.mean(Age)
    ) %>% ungroup()

# Let's address the last remaining passenger (#980).
# After some investigation, I observed that there are only 2 "Ms" in the data
# I will assign them both the same age

data <- data %>%
    group_by(Title) %>%
    mutate(
        Age = impute.mean(Age)
    ) %>% ungroup()
```

### Whats happening to cabin?

```{r}
data %>% group_by(Cabin) %>% tally(sort = TRUE)

# With such a high number of missing values form this column,
# I have decided against including it as a variable to test
```

### Data Wrap-up

Let's wrap-up the data by excluding the columns that are not of interest, and make it all appropriate format.

```{r}
data <- data %>% select(-Ticket,
                -TicketLetter,
                -Cabin,
                -FirstMiddleOther)

data$LastName <- as.numeric(data$LastName)
data$Title <- as.numeric(data$Title)
data$Sex <- as.numeric(data$Sex)
data$Embarked <- as.numeric(data$Embarked)
data$OtherName <- as.numeric(data$OtherName)


summary(data)
```

Write files for supervised learning classificaiton.

```{r}
# Separate the train and test tibbles
sets <- split(data, f = data$tibble)

# create new train tibble
write_csv(x = sets[[2]] %>% select(-tibble), path = "new_train.csv", col_names = TRUE)

# create new test tibble
write_csv(x = sets[[1]] %>% select(-tibble, -Survived), path = "new_test.csv", col_names = TRUE)
```


```{r}
knitr::knit_exit()
```

## Scrap work

Sanity check on all categorical columns

```{r}
s1 <- train %>% 
  ggplot() +
  geom_histogram(aes(Survived), stat = "count") 
  
s2 <- train %>% 
  ggplot() + 
  geom_histogram(aes(Pclass), stat = "count")

s3 <- train %>% 
  ggplot() + 
  geom_histogram(aes(Sex), stat = "count")

s4 <- train %>% 
  ggplot() + 
  geom_histogram(aes(SibSp), stat = "count")

s5 <- train %>% 
  ggplot() + 
  geom_histogram(aes(Parch), stat = "count")

s6 <- train %>% 
  ggplot() + 
  geom_histogram(aes(Embarked), stat = "count")


grid.arrange(s1,s2,s3,s4,s5,s6, ncol=3)
```
